---
title: "Introduction to the farsdata-package"
author: "Florian Handke"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the farsdata-package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This package is primarily built for educational purposes. The package enables one to explore fatal traffic accidents from 2013-2015 using data from the National Highway Traffic Safety Administration (NHTSA) Fatality Analysis Reporting System (FARS).  

## Data

The data in this package come from the National Highway Traffic Safety Administration (NHTSA) Fatality Analysis Reporting System (FARS) data.

```{r sample2013_data}
filename_2013 <- farsdata::make_filename(2013)
farsdata::fars_read(filename_2013) 
```

You can find detail information of the data on [NHTSA FARS Manuals & Documentation page](https://crashstats.nhtsa.dot.gov/#/DocumentTypeList/4). 

## Load FARS Data

Loading the data can be done by using the  `make_filename()` and `fars_read()` functions

### About the Filename

Use the `make_filename` command to find out where data is stored on your machine and/or create a filename to save/load new data.

```{r make_filename}
farsdata::make_filename(2014)
```

### Single Year

Loading data for a signele year can be done by using the function `fars_read_years()` with a single year as input. The only data columns selected are `MONTH` and `year`.  This returns a list of length one, and the first element in the list is the `tbl_df` (the `tidyverse` data frame) listing the month and year for each fatal accident.

```{r single_year}
farsdata::fars_read_years(years = 2014)[[1]]
```

### Multiple Years

Loading multiple years can be done the function `fars_read_years()`.

```{r multiple_years}
farsdata::fars_read_years(years = 2013:2014)
```

## Summarizing FARS Data

The `fars_summarize_years()` function produces a summary of the simple counts of fatalities by month and year: 

```{r summarize_data}
farsdata::fars_summarize_years(2013:2014)
```


## Mapping Fatal Crashes

By assigning a state ID and a year to the function`fars_map_state` a map that state's fatalities with a dot at the fatality location is created.    

State ID numbers can be found [here](https://crashstats.nhtsa.dot.gov/Api/Public/ViewPublication/812315#27). 

```{r mapping_crashes}
library(mapdata)
farsdata::fars_map_state(44, 2014)
farsdata::fars_map_state(8, 2013)
```
